Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-30 manuscript/manuscript.Rmd
PBA edits on July 17 draft

Abstract
--------
Forecasting future states of populations has taken on new urgency as the rate of climate change increases. Traditional plant population models have limited utility in this regard because they are based on detailed demographic data from small, localized plots. These models are difficult to scale up to spatial scales relevant to land managers that require such forecasts to make decisions. To overcome the data limitations of traditional population models, some have proposed population models based on population level, rather than individual level, data that is much easier to collect over broad spatial scales. Using such models violates a central assumption of ecology: individuals respond to weather, not populations to climate. Here, we test whether this assumption is important when forecasting climate change impacts on four perennial grass species in a semi-arid Montana grassland. We parameterized two population models, one based on inidividual-level data with three vital rates and one on population-level data from the same dataset (percent cover of quadrats), and compared their accuracy, precision, and sensitivity to climate. The individual level model was more accurate and precise than the aggregate level model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model tends to "miss" important climate effects from at least one vital rate for each species. We also find that increasing the sample size at the population-level would not necessarily reduce forecast certainty; meaning the only way to reduce uncertainty is to capture unique climate dependence on individual vital rates. It appears there is no short cut to forecasting climate change impacts on plant populations — detailed demographic data is essential. However, forecasts from our individual-level model were very uncertain, so we advocate for a focus on new methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

Building on @Taylor2004, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 demonstrated the application of such methods in a study on arable weeds. In particular, @Queenborough2011 provide empirical evidence that density-structured models are capable of reproducing population dynamics, even if some precision is lost when compared to fully continuous models. The study by @Queenborough2011 included data from 500 fields (4 hectares each) in 49 farms, all collected by two people in 6 weeks. This is far more data from a far greater spatial extent than possible if measuring individual plant demography (in a world of limited time and money, at least). The appeal of density-structured approaches is clear. However, none of these models included environmental covariates.

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact puts us in uneasy proximity to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is an epiphenomenon of demographic processes like survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may difficult resolve in statistical models based on population-level data where demographic processes are not identifiable. If population-level data cannot detect important impacts of climate drivers on populations, then population models built with such data will make poor forecasts.

Here, we ask whether statistical and population models based on aggregated, population-level data can detect climate signals as wells as models based on individual-level data. We used a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. In both models, interannual variation is explained, in part, by climate covariates. We then performed simulations to quantify the sensitivities of species' cover to small perturbations in average precipitation and temperature. We found that population models based on detailed demographic data are more accurate and precise than models based on aggregated data. Our results suggest the popuation-level model is less accurate and precise because important demographic climate signals go undetected. For these species at this location, detailed demographic data appears necessary to make accurate forecasts. A worrying caveat to our work is that forecasts from both models were very uncertain when we considered full process and parameter uncertainty. Even 14 years worth of demographic data may not be sufficient to make useful forecasts.

@@

From 1932 to 1945 individual plants were identified and mapped annually in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were distributed in six pastures, each assigned a grazing treatment of light (1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76 ha/aum) stocking rates (two pastures per treatment). In this analysis we account for potential differences among the grazing treatments, but do not focus on grazing$\times$climate interactions. The annual maps of the quadrats were digitized and the fates of individual plants tracked and extracted using a computer program [@Lauenroth2008; @Chu2014a]. Daily climate data, which we aggregated into climate variables of interest, are available for the duration of the data collection period (1932 - 1945) from the Miles City airport, Wiley Field, 9 km from the study site.

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data we fit models for survival, growth, and recruitment for each species. At the quadrat-level we fit a single regression model for population growth. We describe the statistical models separately since fitting the models required different approaches. All models contain five climate covariates that we chose \emph{a priori}: "water year" precipitation at \emph{t}-1 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$ size interactions. Climate $\times$ size interactions are for climate main effects only, that is we do not include interactions among size and interacting climate effects.

@@

@@

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of including parameter uncertainty in our population models, we saved the final 1,000 iterations from each of the three MCMC chains to be used as randomly drawn values during population simulation. This step alleviates the need to reduce model parameters by model selection since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero. We report the posterior mean, standard deviation, and 95% Bayesian Credible Intervals for every parameter of each model for each species in Appendix B.

### Population models
With the posterior distribution of the vital rate statistical models in hand, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to model populations based on individual level data [cite Ellner and Rees 2006] and an quadrat based version of an individually-based model (Quadrat-Based Model, QBM) to model populations based on quadrat level data. We describe each in turn.

#### Integral projection model
We use an environmentally stochastic IPM [@Rees2009] that includes the random year effects and the climate covariates from the vital rate statistical models. However, for some simulations, we ignore the random year effects so that only the climate effects drive interannual variation. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. So, the density function for any size _v_ at time $t+1$ is

@@

#### Quad-based model
Our quad-based model (QBM) perfectly mirrors its statistical description (Eqs. 8-9). We use the same approach for drawing parameter values as described for the IPM. After drawing the appropriate parameter set, we calculate the mean response (population cover at *t+1* = $x_{t+1}$) according to Eq. 8. We then make a random draw from a [0,1] truncated lognormal distribution with mean equal to $x_{t+1}$ from Eq. 8 and the variance estimate from the fitted model. We can then iterate the model forward by drawing a new parameter set (unique to climate year and MCMC iteration) at each timestep.

@@

### Testing sensitivity to climate covariates
Our main goal in this paper is to see if models based on aggregated data are as sensitive to climate covariates as models based on individual level data. So, with our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model and parameter uncertainty on forecast precision.

@@

Our expectation is that the IPM will produce more accurate and precise forecasts. This could be due to differences in sample sizes leading to larger parameter uncertainty for the QBM, or due to the QBM climate effects being weakly associated with one or more vital rate climate effects at the individual level. To assess the impact of sample size on QBM parameter uncertainty we refit the QBM statistical model (Eqs. 8-9) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats), removing a different randomly selected set of quadrats for each fit. We calculated the standard deviation of climate main effects (pptlag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over replicates within each set of quadrat removals. This allowed us to regress sample size against parameter uncertainty.

To see if the QBM climate effects are correlated, or not, with climate effects for each vital rate model for the IPM, we simply regressed the QBM climate coefficients against each vital rate model's climate coefficients and calculate Pearson's $\rho$. Strong correlations indicate the QBM is capable of detecting climate effects associated with individual vital rates. A weak correlation indicates the QBM "misses" the climate effect on a particular vital rate.

Results
-------
### Comparison of forecast models
The IPM had significantly lower overall error (MAE, mean absolute error) for three species (*B. gracilis*, *H. comata*, *P. smithii*; Table 1). In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (1) lower MAE for three of the four species, (2) statistically similar MAE for the one other species, and (3) considerably more precise forecasts for all species.

### Sensitivity of models to climate
The response of a population to climate change is a result of the aggregate effects of climate on individual vital rates. Since the IPM approach relies on vital rate regressions, we were able to quantify the sensitivity of each vital rate in isolation and in pairwise combinations. Across all species, climate covariates can have opposing effects on different vital rates (Fig. 3). Growth was the most sensitive vital rate for all species, showing a negative response to increased precipitation, and stronger positive response to increased temperature, and a mostly positive response when both climate factors are increased (Fig. 3). *B. gracilis* survival rates were sensitive to temperature, resulting in an increase in plant cover under increased temperature (Fig. 3a). In isolation, recruitment and survival were insensitive to climate factors for *H. comata* (Fig. 3b). Survival and recruitment of *P. smithii* were both sensitive, negatively, to temperature and precipitation (Fig. 3c). *P. secunda* equilibrium cover was sensitive to the climate effects on survival and recruitment, showing a negative effect on both vital rates for increased precipition, but a strong positive effect on survival with increased temperature (Fig. 3d). The climate impact of recruitment on equilibrium cover was negative for precipitation and temperature increases (Fig. 3d). At least two of three vital rates were sensitive to climate for each species (Fig. 3).

### Sources of uncertainty in the QBM
Sample size had a relatively weak effect on QBM climate parameter uncertainty after the number of quadrats exceeded about 10 (Fig. 5). Inverse-gaussian fits show that increasing sample size beyond the number of quadrats we used will results in deminishing returns in terms of parameter certainty (Fig. 5).

@@

### Model forecasts
Forecasts based on 1% climate changes were extremely uncertain when we considered model error and parameter uncertainty (Fig. 6). As expected based on model validation (Table 1), QBM projections were more uncertain than IPM projections for all species except *P. smithiii* (Fig. 6). IPM forecasts for *P. smithiii* were very uncertain due to overcompensating density dependence and nonlinearities recruitment that get amplified by parameter uncertainty (Appendx C). Neither model was capable of making forecasts of proportional cover change distinguishable from 0 (i.e., no change in population state) when we included model error and parameter uncertainty.

@@

Discussion
----------
Population models built using individual-level data allow inference on demographic processes, but their ability to forecast future population states is limited by the spatial extent of the data. Population-level data, like percent cover or discrete density states, are much easier to collect of broad spatial extents, so populations built using such data offer an appealing alternative to traditional population models [@Queenborough2011]. However, at their core, density-structured models rely on individual level data aggregated to a population level metric. This creates a potential problem if such models are to be used in a climate change context because inidividuals respond to climate, not populations [@Clark2012]. Are models based on population level metrics as sensitive to climate as models based on individual level metrics? Do these two types of models produce consistent forecasts? Do we need detailed demographic data to forecast the impacts of climate change? These are the questions we sought to answer here.

Using individual-level and population-level forms of the same dataset, we were able to directly compare a traditional demographic modeling approach to a population model based on percent cover data. Our quad-based model (QBM) is based on percent cover data and so is in the spirit of density-structured models. In terms of each model's forecasting ability, the IPM outperformed the QBM: the IPM was more accurate and precise than the QBM (Table 1). As we suggest in the introduction, this could be due to differences in sample size or the effect of averaging over climate responses unique to individual-level vital rates.

We found no evidence that increasing sample size at the quadrat level would lead to higher precision of climate coefficient estimates (Fig. 4). We did, however, find evidence that the QBM statistical model failed to identify climate dependence for some vital rates (Fig. 5). For no species were climate effects from the QBM strongly correlated with all three vital rates (Fig. 5). @Freckleton2011 acknowledge that averaging over complex stage dependence will lead to poorly specified models. This is analagous to our situation, but instead of averaging over complex life histories, we are averaging over complex climate dependence.

Thus, for each species the QBM is "missing" climate signals associated with at least one vital rate. This leads to inaccurate and imprecise forecasts because the QBM statistical model struggles to explain variation due to climate variables that have positive and negative impacts on different vital rates. When this is the case, as it is for all our species to varying degrees (Fig. 3), statistical models of population-level data will fail. This result further confirms related work on the importance of individual-level data to forecast population responses to exogenous drivers [@Clark2011; @Clark2011b; @Clark2012; @Galvan2014].

These results lead us to conclude that detailed demographic data is necessary to forecast climate change impacts on plant populations when vital rates have unique climate responses. This is unwelcome news since this data is difficult to collect and the models built on such data are of little use to land managers that make decisions at scales beyond that of traditional population models [@Queenborough2011]. While density-structured approaches may fail when climate covariates are considered, there are other alternatives. For example, @Clark2011 use Forest Inventory and Analysis (FIA) data to parameterize a population model with multiple vital rates and climate dependence. Another example are distributed efforts like PlantPopNet (http://plantago.plantpopnet.com) that will allow researchers to estimate variation around climate responses for widespread species by taking advantage of spatial variation in climate [e.g. @Doak2010]. Lastly, we foresee new approaches on the horizon that leverage photo/video of plots and advanced object recognition algorithms [e.g. @Liu2014] to streamline plant mapping and digitizing efforts.

What does all this mean for model forecasts? The answer is more nuanced than one might expect. The typical approach in ecology is to use point estimates of model parameters to project populations forward according to the specified model; usually allowing for some variability around the determinstic process (e.g., @Battin2007; @Jenouvrier2009; @Adler2012). If we follow tradition and calculate the mean response to climate change over many iterations of the model with only model error and temporal variation included, the IPM and the QBM produce opposing forecasts for three of four species (Fig. D1). However, if we introduce parameter uncertainty, the forecasts are actually indistinguishable (Fig. 6). IPM forecasts are generally more precise, as expected from model validation (Table 1). The real story is not that the QBM produces "wrong" forecasts, but that it produces highly uncertain forecasts. Our sample size analysis reveals that this uncertainty is not easily overcome by including more data (Fig. 4). The only way to reduce uncertainty is to model these populations based on individual-level data where climate dependence can be properly identified (Fig. 5). Yet, even the IPM forecasts were highly variable (Fig. 6).

Our goal was not to make any explicit forecast for the future state of these populations based on predicted climate change. But our results highlight the state of affairs in ecology when it comes to forecasting the impacts of climate change. The analysis we conducted here could be considered, with some exceptions of course, at the forefront of ecological forecasting in terms of the statistical approach employed (hierarchical Bayesian), the type of population model we used (stochastic IPM with parameter uncertainty), and the amount of high quality data we had at our disposal (14 years of individual level data). Yet, model predictions proved so uncertain that any forecast, when bounded with model and parameter uncertainty, would be at best not useful and at worst meaningless. For all species, the 90% quantiles of predicted changes in population size overlapped zero; we cannot even predict whether a change in precipitation or temperature will cause populations to increase or decrease. How might we improve on this state of affairs?

First, forecasts could be improved by matching the spatial scale of predictor variables with the spatial scale of observations. One of the major limitations of the models we fit here is that the climate data are at a much larger scale than the individual level observations of plant size. Climate covariates only vary by year, with no spatial variability within years. Thus, even if we fit models to individual-level data, we are missing the key interaction point between weather and individual plants [@Clark2011b] because all observations share the same climate covariates. Demographic studies should be designed with at least plot level measurements of climate related variables (e.g., soil moisture).

Second, accurately detecting climate signals will take even longer time series. Recent theoretical work on detecting climate signals in noisy data suggests that even advanced approaches to parameter fitting like LASSO, functional linear models (splines), and Random Forest models require 20-25 year time series (Teller et al., in review). Alternatively, as we suggest above, Teller et al. (in review) also find that matching the scale of the response and predictors improves estimate precision.

>>>>

Abstract
--------
Forecasting future states of populations has taken on new urgency as the rate of climate change increases. Traditional plant population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. One proposal to overcome the limitations of traditional population models is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. Here, we test the consequences of aggregating individual responses when forecasting climate change impacts on four perennial grass species in a semi-arid Montana grassland. We parameterized two population models, one based on individual-level data (survival, growth and recruitment) and one on population-level data from the same dataset (percent cover of quadrats), and compared their accuracy, precision, and sensitivity to climate. The individual level model was more accurate and precise than the aggregate level model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model tends to "miss" important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. Our analysis indicates that there is no short cut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

Building on @Taylor2004, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 demonstrated the application of such methods in a study on arable weeds. In particular, @Queenborough2011 provide empirical evidence that density-structured models are capable of reproducing population dynamics, even if some precision is lost when compared to fully continuous models. The study by @Queenborough2011 included data from 500 fields (4 hectares each) in 49 farms, all collected by two people in 6 weeks. This is far more data from a far greater spatial extent than would be possible if measuring individual plant demography (in a world of limited time and money). The appeal of density-structured approaches is clear. However, none of these models included environmental covariates.

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes like survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may difficult resolve in statistical models based on population-level data where demographic processes are not identifiable. If population-level data cannot detect important impacts of climate drivers on populations, then population models built with such data will make poor forecasts.

Here, we ask whether statistical and population models based on aggregated, population-level data can predict climate change impacts as wells as models based on individual-level data. We used a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. In both models, interannual variation is explained, in part, by climate covariates. We then performed simulations to quantify the sensitivities of species' cover to small perturbations in average precipitation and temperature. [THE REST OF THIS PARAGRAPH READS TOO MUCH LIKE AN ABSTRACT. CUT OR REDUCE TO ONE SENTENCE? e.g. "individual levels models were clearly better, but, unfortunately, still too uncertain to inform management"] We found that population models based on detailed demographic data are more accurate and precise than models based on aggregated data. Our results suggest the popuation-level model is less accurate and precise because important demographic climate signals go undetected. For these species at this location, detailed demographic data appears necessary to make accurate forecasts. A worrying caveat to our work is that forecasts from both models were very uncertain when we considered full process and parameter uncertainty. Even 14 years worth of demographic data may not be sufficient to make useful forecasts.

@@

From 1932 to 1945 individual plants were identified and mapped annually in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were distributed in six pastures, each assigned a grazing treatment of light (1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76 ha/aum) stocking rates (two pastures per treatment). In this analysis we account for potential differences among the grazing treatments, but do not focus on grazing$\times$climate interactions. The annual maps of the quadrats were digitized and the fates of individual plants tracked and extracted using a computer program [@Lauenroth2008; @Chu2014a]. Daily climate data are available for the duration of the data collection period (1932 - 1945) from the Miles City airport, Wiley Field, 9 km from the study site.

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit models for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately since fitting the models required different approaches. All models contain five climate covariates that we chose \emph{a priori}: "water year" precipitation at \emph{t}-1 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$ size interactions. Climate $\times$ size interactions are for climate main effects only; we do not include interactions between size and pairs of interacting climate effects.

@@

?Is sigma_j a constant, or does it depend on other parameters or state variables?

@@

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recognize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of including parameter uncertainty in our population models, we saved the final 1,000 iterations from each of the three MCMC chains to be used as randomly drawn values during population simulation. This step alleviates the need to reduce model parameters by model selection since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero. We report the posterior mean, standard deviation, and 95% Bayesian Credible Intervals for every parameter of each model for each species in Appendix B.

### Population models
With the posterior distribution of the vital rate statistical models in hand, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to model populations based on individual level data [cite Ellner and Rees 2006] and a quadrat-based version of an individually-based model (Quadrat-Based Model, QBM) to model populations based on quadrat level data. We describe each in turn.

#### Integral projection model
We use an environmentally stochastic IPM [@Rees2009] that includes the random year effects and the climate covariates from the vital rate statistical models. However, for some simulations, we ignore the random year effects so that only the climate effects drive interannual variation \todo{Add spatial structure mention. Maybe you can specify when we use random yrs effects, when not. Just above line 191 I found myself wondering when you do use them }. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size _v_ at time $t+1$ is

@@

#### Quad-based model
Our quad-based model (QBM) perfectly mirrors its statistical description (Eqs. 8-9). We use the same approach for drawing parameter values as described for the IPM. After drawing the appropriate parameter set, we calculate the mean response (population cover at *t+1* = $x_{t+1}$) according to Eq. 8. We then make a random draw from a [0,1] truncated lognormal distribution with mean equal to $x_{t+1}$ from Eq. 8 and the variance estimate from the fitted model. We can then project the model forward by drawing a new parameter set (unique to climate year and MCMC iteration) at each timestep.

@@

### Testing sensitivity to climate covariates
Our main goal in this paper \todo{is this our main goal? it's not quite the same as making good predictions. It is a goal...} is to see if models based on aggregated data are as sensitive to climate covariates as models based on individual level data. So, with our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model and parameter uncertainty on forecast precision.

@@

Our expectation is that the IPM will produce more accurate and precise forecasts. This could be due to 1) the smaller  sample size of the quadrat level data sets compared to the individual level data sets, leading to larger parameter uncertainty for the QBM, or 2) the QBM climate effects being weakly associated with one or more vital rate climate effects at the individual level. To assess the impact of sample size on QBM parameter uncertainty we refit the QBM statistical model (Eqs. 8-9) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats), removing a different randomly selected set of quadrats for each fit. We calculated the standard deviation of climate main effects (pptlag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over replicates within each set of quadrat removals. This allowed us to regress sample size against parameter uncertainty.

To see if the QBM climate effects are correlated, or not, with climate effects for each vital rate model for the IPM, we simply regressed the QBM climate coefficients against each vital rate model's climate coefficients and calculated Pearson's $\rho$. Strong correlations indicate the QBM is capable of detecting climate effects associated with individual vital rates. A weak correlation indicates the QBM "misses" the climate effect on a particular vital rate.

Results
-------
### Comparison of forecast models
The IPM had significantly lower overall error (MAE, mean absolute error) for three species (*B. gracilis*, *H. comata*, *P. smithii*; Table 1). \todo{Perhaps mention the PASM recruitment*climate uncertainty here? to explain why this species is different?} In no case did the QBM significantly outperform the IPM (Table 1). The IPM was consistently more precise than the QBM, with lower distances between the 90% quantiles across all species (Table 1). In general the IPM outperformed the QBM because it had (1) lower MAE for three of the four species, (2) statistically similar MAE for the one other species, and (3) considerably more precise forecasts for all species.

### Sensitivity of models to climate
The response of a population to climate change is a result of the aggregate effects of climate on individual vital rates. Since the IPM approach relies on vital rate regressions, we were able to quantify the sensitivity of each vital rate in isolation and in pairwise combinations. Across all species, climate covariates can have opposing effects on different vital rates (Fig. 3). Growth was the most sensitive vital rate for all species, showing a negative response to increased precipitation, and stronger positive response to increased temperature, and a mostly positive response when both climate factors are increased (Fig. 3). *B. gracilis* survival rates were sensitive to temperature, resulting in an increase in plant cover under increased temperature (Fig. 3a). In isolation, recruitment and survival were insensitive to climate factors for *H. comata* (Fig. 3b). Survival and recruitment of *P. smithii* were both sensitive, negatively, to temperature and precipitation (Fig. 3c). *P. secunda* equilibrium cover was sensitive to the climate effects on survival and recruitment, showing a negative effect on both vital rates for increased precipition, but a strong positive effect on survival with increased temperature (Fig. 3d). Equilibrium cover responded negatively when increased precipitation and temperature affect recruitment  (Fig. 3d). At least two of three vital rates were sensitive to climate for each species (Fig. 3).

### Sources of uncertainty in the QBM
Sample size had a relatively weak effect on QBM climate parameter uncertainty after the number of quadrats used in fitting exceeded about 10 (Fig. 5). Inverse-gaussian fits show that increasing sample size beyond the number of quadrats we used would result in diminishing returns in terms of parameter certainty (Fig. 5).

@@

### Model forecasts
Forecasts based on 1% climate changes were extremely uncertain when we considered model error and parameter uncertainty (Fig. 6). \todo{I thought you were going to show the no-parameter-uncertainty forecasts in appendix?.} As expected based on model validation (Table 1), QBM projections were more uncertain than IPM projections for all species except *P. smithiii* (Fig. 6). IPM forecasts for *P. smithiii* were very uncertain due to a very high instrinsic rate of recruitment combined with uncertainty in climate coefficients which lead to high recruitment boom years and subsequent busts when young plants suffer high mortality (Appendx C). When we included model error and parameter uncertainty, forecast changes in proportional cover always spanned a wide range of negative to positive values. In other words, neither model could predict whether a climate perturbation would increase or decrease equilibrium population size.

@@

Discussion
----------
Population models built using individual-level data allow inference on demographic processes, but they can only forecast future population states across the (typically limited) spatial extent of the observations. Population-level data are much easier to collect across broad spatial extents, so models built using such data offer an appealing alternative to traditional population models [@Queenborough2011]. However, density-structured models rely on the aggregation of individual level data. This creates a potential problem if such models are to be used in a climate change context because it is individuals, not populations, which respond to climate [@Clark2012]. Are models based on population level metrics as sensitive to climate as models based on individual level metrics? Do these two types of models produce consistent forecasts? Do we need detailed demographic data to forecast the impacts of climate change?

### The importance of demographic data

Our comparison of a traditional, demographic population model (the IPM) with a model inspired by density-structured models (the QBM) showed that the IPM outperformed the QBM: the IPM was more accurate and precise than the QBM in out-of-sample cross validation (Table 1). The superiority of the IPM could reflect either differences in sample size or the effect of averaging over unique effects of climate on each individual-level vital rate. Although increasing sample size of quadrat percent-cover observations would be easy to do in the field, we little evidence that it would lead to higher precision of climate coefficient estimates (Fig. 4). We did, however, find evidence that the QBM statistical model failed to identify climate dependence for some vital rates (Fig. 5). For no species were climate effects from the QBM strongly correlated with all three vital rates (Fig. 5). @Freckleton2011 acknowledge that averaging over complex stage dependence will lead to poorly specified models. This is analagous to our situation, but instead of averaging over complex life histories, we are averaging over complex climate dependence.

Our interpretation is that the QBM is "missing" climate signals associated with at least one vital rate for each species. This leads to inaccurate and imprecise forecasts because the QBM statistical model struggles to explain variation due to climate variables that have positive and negative impacts on different vital rates. When this is the case, as it is for all our species to varying degrees (Fig. 3), statistical models of population-level data will fail. Our result is consistent with related work on the importance of individual-level data to forecast population responses to exogenous drivers [@Clark2011; @Clark2011b; @Clark2012; @Galvan2014].

Detailed demographic data appears to be necessary to forecast climate change impacts on plant populations when vital rates have unique climate responses. How then can we build models to make forecasts for the landscape and regional scales beyond the scope of traditional population models [@Queenborough2011]? There are alternatives to density-structured models. For example, @Clark2011 use Forest Inventory and Analysis (FIA) data to parameterize a population model with multiple vital rates and climate dependence. Distributed efforts such as PlantPopNet (http://plantago.plantpopnet.com) will allow researchers to estimate variation around climate responses for widespread species by taking advantage of spatial variation in climate [e.g. @Doak2010]. Finally, new approaches on the horizon that leverage photo/video of plots and advanced object recognition algorithms [e.g. @Liu2014] will increase the efficiency of plant mapping and digitizing efforts.

### The challenge of uncertainty

An important, but unexpected, result of our analysis was the great uncertainty in forecasts, even for our best model. The typical approach in ecology is to use point estimates of model parameters to project populations forward according to the specified model, usually allowing for some variability around the determinstic process (e.g., @Battin2007; @Jenouvrier2009; @Adler2012). If we follow tradition and calculate the mean response to climate perturbation with only model error and interannual variation included, the IPM and the QBM produce opposing forecasts for three of four species (Fig. D1). It would be tempting to interpret this inconsistency as further evidence for the superiority of the IPM. However, if we introduce parameter uncertainty, the forecasts are actually indistinguishable (Fig. 6), though the IPM projections are generally more precise (consistent with our cross-validation results). The real story is that both models produce highly uncertain forecasts. For all species, the 90% quantiles of predicted changes in population size overlapped zero; we cannot even predict whether a change in precipitation or temperature will cause populations to increase or decrease. \todo{I think we should look at larger perturbations. How large a perturbation do we need to get a forecast that doesn't span 0?}

Our results highlight the state of affairs in ecology when it comes to forecasting the impacts of climate change. The analysis we conducted here could be considered at the forefront of ecological forecasting with repsect to the statistical approach employed (hierarchical Bayesian), the type of population model we used (density-dependent, stochastic IPM with parameter uncertainty), and the amount of high quality data we had at our disposal (14 years of individual level data). Yet, model predictions proved so uncertain that any forecast, when bounded with model and parameter uncertainty, would be at best not useful and at worst meaningless.

How might we improve on this state of affairs? First, forecasts could be improved by matching the spatial scale of predictor variables with the spatial scale of observations. One of the major limitations of the models we fit here is that the climate data are at a much larger scale than the individual level observations of plant size. Climate covariates only vary by year, with no spatial variability within years. Thus, even if we fit models to individual-level data, we are missing the key interaction point between weather and individual plants [@Clark2011b] because all observations share the same climate covariates. Demographic studies should be designed with at least plot-level measurements of climate related variables (e.g., soil moisture). Second, accurately detecting climate signals will take even longer time series. Recent theoretical work on detecting climate signals in noisy data suggests that even advanced approaches to parameter fitting require 20-25 year time series (Teller et al., in review). Third, ecologists need a stronger commitment to reporting uncertainty. Although most modeling studies explicitly consider model uncertainty, parameter uncertainty is often ignored. In some cases this is because the most convenient statistical methods make it difficult to propogate parameter uncertainty. Yet even Bayesian approaches that allow integration of model fitting and forecasting [@Hobbs2015] are not simple when using modeling approaches like integral projection models that separate the model fitting and simulation stages [@Rees2009]. However, as we have done here, it is still possible to include parameter uncertainty by drawing parameter values from MCMC iterations, taking care to draw all parameters from the same chain and iteration to account for their correlations. Only by being honest about our forecasts can we begin to produce better ones, and forecasts reported without parameter error are disingenuous. In many cases, ignoring parameter error is justifiable when investigating basic demographic mechanisms, but it is not justifiable when forecasting is the goal.
